### 2020-12-24 latex/acrobot.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose the orbit \(x(\R)\) is confined within 
\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
zone.
Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
implies:
\begin{itemize}
    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
        invariant.
    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
\end{itemize}
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
There are two possibilities: either \(E_\pi\) itself is invariant, or the
largest invariant subset of \(E_\pi\) is \(\{(\pi,0)\}\).
To rule out the first possibility, take the derivative of \(E(q_u,p_u)\) at the
\(p_u\)-axis to get
\[
    \dot{E}(0,p_u) = -\frac{g}{5l} \sin(q_a)p_u
    .
\]
This is non-zero everywhere except at the origin, which means it is non-zero on
\(E_\pi\). 
This means \(E_\pi\) is not invariant, so the positive limit set of \(x(t)\)
must be the set \(\{(\pi,0)\}\).
Hence, \(x(t)\) converges to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

@@

\section{Experimental Results}
In this section, we test our VNHC on the physical acrobot built by
\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
Since it is not a simple acrobot, this experiment will test whether VNHCs are
robust to model mismatch.
As we will see, our VNHC does inject energy on the general acrobot model.

>>>>

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose by way of contradiction that the orbit \(x(\R)\) is confined within 
\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
zone.
Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
implies:
\begin{itemize}
    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
        invariant.
    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
\end{itemize}
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
There are two possibilities: either \(E_\pi\) itself is invariant, or the
largest invariant subset of \(E_\pi\) is \(\{(\pi,0)\}\).
To rule out the first possibility, take the derivative of \(E(q_u,p_u)\) at the
\(p_u\)-axis to get
\[
    \dot{E}(0,p_u) = -\frac{g}{5l} \sin(q_a)p_u
    .
\]
This is non-zero everywhere except at the origin, which means it is non-zero on
\(E_\pi\). 
This means \(E_\pi\) is not invariant, so the positive limit set of \(x(t)\)
must be the set \(\{(\pi,0)\}\).
Hence, \(x(t)\) converges to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

@@

\section{Experimental Results}
In this section, we test our VNHC on the physical acrobot built by
Xingbo Wang \cite{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
This acrobot is not simple, that is, the torso and leg rods have unequal mass
and differ in length. 
Furthermore, the centers of mass are not located at the tip of each rod.
These experiments will therefore test whether VNHCs are robust to model
mismatch, since our VNHC is only proven to inject energy on simple acrobots.
As we will see, our VNHC does inject energy on the general acrobot model.

### 2020-12-24 latex/introduction.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

The Estonian sport of ``Kiiking" (Figure \ref{fig:kiiking}) is the adult
equivalent of a child playing on a standing swing, though here
an athlete attempts to make a giant solid steel swing gain momentum.
Naturally, the kiiker begins by pushing off the ground and proceeds to alternate
between standing and squatting. 
This allows them to swing higher and faster over time, until they have reached
their express goal of spinning at high speed around the bar.

@@

The modern concept of virtual nonholonomic constraints was described by
\citet{nhvc_dynamic_walking} in 2015, though there are references to more
primitive versions going back as early as the the year 2000
\cite{vnhc_human_robot_cooperation}.
Virtual nonholonomic constraints have been of most notable use in
bipedal locomotion \cite{nhvc_incline_walking,output_nhvc_bipedal_control},
where they have shown marked improvements in the robustness of walking gaits
when compared to previous control technqiues.
They have also been used for error-reduction in time-delayed teleoperation
\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
All these research areas use different formulations of virtual nonholonomic constraints;
\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
zero dynamics for hybrid mechanical systems under virtual nonholonomic
constraints, but to the best of our knowledge no one has attempted to generalize
the concept to abstract mechanical systems.

>>>>

The Estonian sport of \textit{Kiiking} (Figure \ref{fig:kiiking}) is the adult
equivalent of a child playing on a standing swing, though here
an athlete attempts to make a giant solid steel swing gain momentum.
Naturally, the kiiker begins by pushing off the ground and proceeds to alternate
between standing and squatting. 
This allows them to swing higher and faster over time, until they have reached
their express goal of spinning at high speed around the bar.

@@

The modern concept of virtual nonholonomic constraints was described by
\citet{nhvc_dynamic_walking} in 2015, though there are references to more
primitive versions going back as early as the the year 2000
\cite{vnhc_human_robot_cooperation}.
Virtual nonholonomic constraints have been of most notable use in
bipedal locomotion 
\cite{nhvc_gait_optimization,output_nhvc_bipedal_control},
where they have shown marked improvements in the robustness of walking gaits
when compared to previous control techniques.
They have also been used for error-reduction in time-delayed teleoperation
\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
zero dynamics for hybrid mechanical systems under virtual nonholonomic
constraints.
In particular, they study a class of constraints formed by taking an affine
sum of holonomic functions with nonholonomic B\'{e}zier polynomials.
They widen this class of constraints in \cite{nhvc_incline_walking} to improve
bipedal locomotion on variable-slope terrain, and generate constraints through
a specialized optimization routine.

Each of these research areas use different formulations of virtual
nonholonomic constraints.
Furthermore, the most recent literature surrounding virtual nonholonomic constraints
has focused on enforcing stable periodic orbits for hybrid mechanical systems
with impacts.
To the best of our knowledge, no one has attempted to create a single unified
framework of virtual nonholonomic constraints for general mechanical systems.
Moreover, no one has studied how virtual nonholonomic constraints can be used for energy
injection, where periodic motion is not desirable.
This thesis attempts to fill these gaps in the research.

### 2020-12-24 latex/vnhcs.tex
Added more detail to lit review and made edits as per Manfredi's suggestions

\begin{cor}\label{cor:2d-zero-dynamics}
    Let \(\mathcal{H}\) be the system ~\eqref{eqn:simple-hamiltonian} with degree
    of underactuation one.
    Suppose \(\mathcal{H}\) satisfies Assumption \ref{assm:inertially-actuated}.
    Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form
    \(h(q,p_u) = q_a - f(q_u,p_u)\),
    where \(f\) is a suitably defined \(C^2\) function.
    Then \(dh_q = \begin{bmatrix} -\partial_{q_u}f & \Id{(n-1)}
    \end{bmatrix}\).
    Defining \(e_1 := (1,0,\ldots,0)\tpose \in \R^n\), the actuated momentum is
    \begin{equation}\label{eqn:g-qupu}
        p_a = -\left.\left(dh_q \Minv(q)
        \begin{bmatrix}
            \Zmat{1\times (n-1)}\\
            I_{(n-1)}
        \end{bmatrix}\right)\inv 
        \left(\partial_{p_u}f \partial_{q_u}V + dh_q \Minv(q)e_1 p_u\right) 
            \right|_{q_a =f(q_u,p_u)}
        .
    \end{equation}
    Since \(q_u \in \Rt{T_u}\) for some \(T_u \in ]0,\infty]\) and \(p_u \in \R\),
    the orbit \((q_u(t),p_u(t))\) traces out
    a curve on the 2D constraint manifold 
    \(\Gamma \simeq \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
\end{cor}

@@

%%/========== /Virtual Nonholonomic Constraints ==========/% 
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :

>>>>

\begin{cor}\label{cor:2d-zero-dynamics}
    Let \(\mathcal{H}\) be the system ~\eqref{eqn:simple-hamiltonian} with degree
    of underactuation one.
    Suppose \(\mathcal{H}\) satisfies Assumption \ref{assm:inertially-actuated}.
    Let \(h(q,p_u) = 0\) be a regular VNHC of order \((n-1)\) of the form
    \(h(q,p_u) = q_a - f(q_u,p_u)\),
    where \(f\) is a suitably defined \(C^2\) function.
    Then \(dh_q = \begin{bmatrix} -\partial_{q_u}f & \Id{(n-1)}
    \end{bmatrix}\).
    Defining \(e_1 := (1,0,\ldots,0)\tpose \in \R^n\), the actuated momentum is
    \begin{equation}\label{eqn:g-qupu}
        p_a = -\left.\left(dh_q \Minv(q)
        \begin{bmatrix}
            \Zmat{1\times (n-1)}\\
            I_{(n-1)}
        \end{bmatrix}\right)\inv 
        \left(\partial_{p_u}f \partial_{q_u}V + dh_q \Minv(q)e_1 p_u\right) 
            \right|_{q_a =f(q_u,p_u)}
        .
    \end{equation}
    Since \(q_u \in \Rt{T_u}\) for some \(T_u \in ]0,\infty]\) and \(p_u \in \R\),
    the orbit \((q_u(t),p_u(t))\) traces out
    a curve on the constraint manifold 
    \(\Gamma \simeq \Rt{T_u} \times \R\) which we call the \((q_u,p_u)\)-plane.
\end{cor}

@@

%/========== /Virtual Nonholonomic Constraints ==========/% 
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :
